---
title: "Statistical learning: MDSR Chapter 8 (STAT495)"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "October 31, 2016"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
library(mdsr)   # Load additional packages here 

# Some customization.  You can alter or delete as desired (if you know what you are doing).
trellis.par.set(theme=theme.mosaic()) # change default color scheme for lattice
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits=4)
```

### HW #5 sample solution

### Project feedback

### MSA

#### What's working well

- R and applications; real work experiences
- breadth of topics
- lecturing is getting better (feel safer)
- collaboration plus individual work

#### What's working not quite as well

- too many deliverables, need more time!
- Wikipedia: how does it fit in?
- CASI on test (surprise!)

#### How can Nick help you master learning outcomes for the course?

- more time!
- too many little assignments
- spark and bigger data
- non-class day office hours
- more info on factor analysis and ridge regression

### Learning from Data

Flight delays have serious costs in terms of time and money. The `nycflights13` data set has a table `flights` that contains a continuous `arr_delay` that indicates the deviation between the actual and scheduled arrival time. Use this variable to create a dichotomous indicator `late` which is true if the arrival delay is greater than 15 minutes.  For each of the following models:
  \begin{enumerate}
    \item Build at least three classifiers for `late`
    \item Compare their effectiveness on the `nycflights13` training data
    \item Make an appropriate visualization of the model
    \item Interpret the results. What have you learned about flight delays?
  \end{enumerate}

My advice: for now keep it simple!

```{r}
library(nycflights13)
set.seed(1999)
flights <- flights %>%
  mutate(trainprob = runif(nrow(flights)),
         late = ifelse(arr_delay > 15, "late", "on time")
  ) # your recoding and the like go here
train <- filter(flights, trainprob < 0.8)
test <- filter(flights, trainprob >= 0.8)
dim(train)
dim(test)
tally(~ late, format="prop", data=train)
```

```{r}

